Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change

### 2018-09-29 LatexNotes/Notes.tex
NM

% Experimental parameters
\newcommand{\lamm}{\lambda^{-}}
\newcommand{\lamp}{\lambda^{+}}
\newcommand{\hm}{h^{-}}
\newcommand{\hp}{h^{+}}
\newcommand{\etastd}{\eta_{\text std}}
% Theory for truncated KdV and stat mech
\newcommand{\uhat}{\hat{u}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\Real}{\text{Re}}
\newcommand{\Fspace}{\mathscr{F}_{\Lambda}}
\newcommand{\Proj}{P_{\Lambda}}
\newcommand{\sumk}{\sum_{k=1}^{\Lambda}}
\newcommand{\sumn}{\sum_{n=1}^{N}}
\newcommand{\uDir}{u_{\text{Dir}}}
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\ampp}{\mathcal{A}}

@@

There are a few important things to remember when we reformulate this equation:
\begin{enumerate}
\item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
\item We want to normalize the standard deviation (or energy) of $\eta$ and therefore want to keep Johnson's choice to scale surface displacement on $a = \eps h_0$.
\item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change. We have to do this so that the assumption of spatial periodicity will be reasonably satisfied on both sides.
\end{enumerate}
To achieve 3, we can use the characteristic wavelength from shallow water theory,
\begin{equation}
\lambda_c = {c}/{f_c} = \sqrt{g d}/f_c
\end{equation}
where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed. It is tempting to simply set $f_c$ to the peak forcing frequency $f_p$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$. Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments). We therefore choose
\begin{equation}
\lambda_c = M \lambda_p = M \sqrt{gd} / f_p
\end{equation}
where $M$ is an integer (on the order of 10 or so). Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient. Thus, the main change we need to make is to introduce the dimensionless variable
\begin{equation}
\zeta = \frac{2\pi}{\lambda_c} (\bar{x} - c \bar{t}) 
= \frac{2 \pi \delta}{M \eps^{1/2}} \, \xi
\end{equation}
where
\begin{equation}
\delta = h_0 / \lambda_0 = f_p \sqrt{h_0/g}
\end{equation}
To clean things up a little, lets also introduce a slightly different long-space coordinate that uses the upstream wavelength $\lambda_0$ as a characteristic scale instead of $h_0$,
\begin{equation}
s = \frac{2 \pi \beta}{M \lambda_0} \bar{x} = 2 \pi \beta M^{-1} \delta \eps^{-3/2} X
\end{equation}
where $\beta$ is a small parameter that scale like $\eps$ or $\delta^2$ for example.

@@

Equation \eqref{varKdVugly} is essentially in the form we want to use, except that the notation is bad. In particular, we will want to reserve the symbol $H$ for the Hamiltonian. So lets introduce some new variables
\begin{align}
& u = D^{1/4} \, \bar{\eta}/a \\
& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
& s = 2 \pi \beta M^{-1} \lambda_0^{-1} \bar{x}
\end{align}
Above, the original physical variables are denoted with a bar. Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}[h]%[htbp]
\begin{center}
\caption{Table of parameters}
\label{paramtable}
\begin{tabular}{l l l l}
\hline \multicolumn{4} { c }{Parameters that are constant throughout the domain} \\
\hline Description & Notation and definition & Units & Value in experiments \\
\hline
Peak forcing frequency		& $f_p$						& frequency	& 2 Hz \\
Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2}$		& length		& 0.03--0.3 cm \\
Upstream wavelength			& $\lambda_0 = \sqrt{g h_0}/f_p$	& length		& 56 cm \\
Amplitude-to-depth ratio		& $\eps = a / h_0$				& dimensionless	& 0.0024--0.024 \\
Depth-to-wavelength ratio		& $\delta = h_0 / \lambda_0$		& dimensionless		& 0.22 \\
Amplitude parameter			& $\ampp = \eps \delta^{-2} = ag h_0^{-2} f_p^{-2}$	
& dimensionless		& 0.05--0.5\\
Catch-all parameter		& $\mu = 4 \pi^2 / (9 M^2)$			& dimensionless		& 0.27 if $M=4$ \\
\hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
\hline Description & Notation and definition & Value in experiments \\
\hline
Local depth			& $d$					& length \\
Local wavespeed		& $c = \sqrt{gd}$			& speed \\
Local peak wavelength	& $\lambda_p = \sqrt{gd}/f_p$	& length \\
Characteristic wavelength	& $\lambda_c = M \lambda_p$	& length \\
Depth ratio			& $D = d/h_0$				& dimensionless
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

In Table \ref{paramtable} we provide a list of all the relevant parameters needed to convert these dimensional variables to physical ones. Note that the wavelength $\lambda_0$ is predicted from shallow-water theory and so it differs somewhat from the value in Table \ref{exptable} obtained from the dispersion relation. In terms of the new variables,  Eq.~\eqref{varKdVugly} becomes
\begin{align}
\label{varKdV1}
2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
+ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} \, u_{xxx} = 0
\end{align}

@@

Then the Hamiltonian structure is given by
\begin{align}
\label{Hamiltonian}
& H^- = \ampp H_3 - \mu H_2 \\
& H^+ = \ampp D_0^{-7/4} H_3 - \mu D_0^{1/2} H_2
\end{align}

>>>>

% Experimental parameters
\newcommand{\lamm}{\lambda^{-}}
\newcommand{\lamp}{\lambda^{+}}
\newcommand{\hm}{h^{-}}
\newcommand{\hp}{h^{+}}
\newcommand{\etastd}{\eta_{\text std}}
% Theory for truncated KdV and stat mech
\newcommand{\uhat}{\hat{u}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\Real}{\text{Re}}
\newcommand{\Fspace}{\mathscr{F}_{\Lambda}}
\newcommand{\Proj}{P_{\Lambda}}
\newcommand{\sumk}{\sum_{k=1}^{\Lambda}}
\newcommand{\sumn}{\sum_{n=1}^{N}}
\newcommand{\uDir}{u_{\text{Dir}}}
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\ampp}{\mathcal{A}}
\newcommand{\lamfac}{N}

@@

There are a few important things to remember when we reformulate this equation:
\begin{enumerate}
\item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
\item We want to normalize the standard deviation (or energy) of $\eta$ and therefore want to keep Johnson's choice to scale surface displacement on $a = \eps h_0$.
\item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change. We have to do this so that the assumption of spatial periodicity will be reasonably satisfied on both sides.
\end{enumerate}
To achieve 3, we can use the characteristic wavelength from shallow water theory,
\begin{equation}
\lambda_c = {c}/{f_c} = \sqrt{g d}/f_c
\end{equation}
where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed. It is tempting to simply set $f_c$ to the peak forcing frequency $f_p$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$. Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments). We therefore choose
\begin{equation}
\lambda_c = \lamfac \lambda_p = \lamfac \sqrt{gd} / f_p
\end{equation}
where $\lamfac$ is an integer (on the order of 10 or so). Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient. Thus, the main change we need to make is to introduce the dimensionless variable
\begin{equation}
\zeta = \frac{2\pi}{\lambda_c} (\bar{x} - c \bar{t}) 
= \frac{2 \pi \delta}{\lamfac \eps^{1/2}} \, \xi
\end{equation}
where
\begin{equation}
\delta = h_0 / \lambda_0 = f_p \sqrt{h_0/g}
\end{equation}
To clean things up a little, lets also introduce a slightly different long-space coordinate that uses the upstream wavelength $\lambda_0$ as a characteristic scale instead of $h_0$,
\begin{equation}
s = \frac{2 \pi \beta}{\lamfac \lambda_0} \bar{x} = 2 \pi \beta \lamfac^{-1} \delta \eps^{-3/2} X
\end{equation}
where $\beta$ is a small parameter that scale like $\eps$ or $\delta^2$ for example.

@@

Equation \eqref{varKdVugly} is essentially in the form we want to use, except that the notation is bad. In particular, we will want to reserve the symbol $H$ for the Hamiltonian. So lets introduce some new variables
\begin{align}
& u = D^{1/4} \, \bar{\eta}/a \\
& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
& s = 2 \pi \beta \lamfac^{-1} \lambda_0^{-1} \bar{x}
\end{align}
Above, the original physical variables are denoted with a bar. Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}[h]%[htbp]
\begin{center}
\caption{Table of parameters}
\label{paramtable}
\begin{tabular}{l l l l}
\hline \multicolumn{4} { c }{Parameters that are constant throughout the domain} \\
\hline Description & Notation and definition & Units & Value in experiments \\
\hline
Peak forcing frequency		& $f_p$						& frequency	& 2 Hz \\
Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2}$		& length		& 0.03--0.3 cm \\
Upstream wavelength			& $\lambda_0 = \sqrt{g h_0}/f_p$	& length		& 56 cm \\
Amplitude-to-depth ratio		& $\eps = a / h_0$				& dimensionless	& 0.0024--0.024 \\
Depth-to-wavelength ratio		& $\delta = h_0 / \lambda_0$		& dimensionless		& 0.22 \\
Amplitude parameter			& $\ampp = \eps \delta^{-2} = ag h_0^{-2} f_p^{-2}$	
& dimensionless		& 0.05--0.5\\
Catch-all parameter		& $\mu = 4 \pi^2 / (9 M^2)$			& dimensionless		& 0.27 if $M=4$ \\
\hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
\hline Description & Notation and definition & Units \\
\hline
Local depth			& $d$					& length \\
Local wavespeed		& $c = \sqrt{gd}$			& speed \\
Local peak wavelength	& $\lambda_p = \sqrt{gd}/f_p$	& length \\
Characteristic wavelength	& $\lambda_c = M \lambda_p$	& length \\
Depth ratio			& $D = d/h_0$				& dimensionless
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

In Table \ref{paramtable} we provide a list of all the relevant parameters needed to convert these dimensional variables to physical ones. Note that the wavelength $\lambda_0$ is predicted from shallow-water theory and so it differs somewhat from the value in Table \ref{exptable} obtained from the dispersion relation. In terms of the new variables,  Eq.~\eqref{varKdVugly} becomes
\begin{align}
\label{varKdV1}
2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
+ \frac{4 \pi^2 \delta^2}{3 \lamfac^2} D^{1/2} \, u_{xxx} = 0
\end{align}

@@

\subsection{Hamiltonian structure}
Then the Hamiltonian structure is given by
\begin{align}
\label{Hamiltonian}
& H^- = \ampp H_3 - \mu H_2 \\
& H^+ = \ampp D_0^{-7/4} H_3 - \mu D_0^{1/2} H_2
\end{align}
